---
title: "Bike Share Analysis"
output: html_notebook
---


```{r}

file_names_bike_share= dir("C:\\Users\\dsmit\\OneDrive\\Documents\\MDC Data Analytics\\Capstone\\Bike_Share_First-1.Rmd", full.names=TRUE)

```


```{r}

library ("tidyverse")

```


```{r}

library("readxl")

```


```{r}

bike_share_raw_Oct22Feb23 <- do.call(rbind,lapply(file_names_bike_share,read_excel))

```


Take a look at the data frame:

```{r}

str(bike_share_raw_Oct22Feb23)

```

### Are there missing values?

```{r}
sum (complete.cases(bike_share_raw_Oct22Feb23))
```


Get a count of missing values per column (alternative 1):

```{r}

sapply(bike_share_raw_Oct22Feb23, function(x) {sum(is.na(x))})

```


Get a count of missing values per column (alternative 2):

```{r}

bike_share_raw_Oct22Feb23 %>% select(everything()) %>%summarise_all(funs(sum(is.na(.))))

```


How to handle the missing values?

a) Can we find a way of finding and entering the missing values?
b) Remove rows and find new rows without missing values IF sample size is a concerned
c) Remove rows and use the reduced data frame IF sample size is not a concerned(THIS WILL BE OUR APPROACH)

Other methods to handle missing values are:
Mean imputation (discuss)
Model imputation (discuss)
Regression imputation (discuss)

Are any of these last three methods applicable here?


```{r}

bike_share_Oct22Feb23 <- bike_share_raw_Oct22Feb23[complete.cases(bike_share_raw_Oct22Feb23), ]

```


Checking that the new data frame has no missing values:

```{r}
nrow(bike_share_Oct22Feb23)
sum (complete.cases(bike_share_Oct22Feb23))

```



### Check for possible incorrect values in the columns


```{r}
str(bike_share_Oct22Feb23)
```

Any strange/erroneous values in the qualitative variables?

Check the columns rideable_type and member_casual.

ASK STUDENTS: What am I trying to accomplish with the following code?

```{r}

 table (bike_share_Oct22Feb23$rideable_type)
 table (bike_share_Oct22Feb23$member_casual)

```

I do not think we need to worry about strange values in ride_id. What do you think? And, what would it mean to check for strange values in ride_id?


Are we going to use started_at, ended_at, and the latitude and longitude columns?


Check ride_length for erroneous data:

```{r}

summary(bike_share_Oct22Feb23$ride_length)

```

Something is wrong, the values in ride_lenght were supposed to be in the format hh:mm:ss, but we are getting a date value as well.

```{r}

head (bike_share_Oct22Feb23$ride_length, 10)

```

I am going to solve this issue using a long and probably not the most efficient approach. You can also try using R functions to handle dates such as those part of the "lubridate" package.


```{r}

ride_length_temp= separate (bike_share_Oct22Feb23, ride_length, into= c("date_ride", "time_ride"), sep=" ")$time_ride

```

```{r}
head (ride_length_temp, 10)
```

```{r}

mode(ride_length_temp)

```

Now, grab the hours and minutes (ignore seconds? Yes, I think it is safe to do so) and convert them to an integer. Then, record the ride lenght in minutes (i.e., length= how many minutes the ride lasted)


Getting hours:

```{r}

ride_length_temp_hours= substr(ride_length_temp, start = 1, stop = 2)

```


```{r}

table (ride_length_temp_hours)

```

```{r}

ride_length_temp_hours_int = as.integer (ride_length_temp_hours)
  
```


```{r}
# Sanity check

table (ride_length_temp_hours_int)

```

Getting minutes

```{r}

ride_length_temp_mins= substr(ride_length_temp, start = 4, stop = 5)

```


```{r}
table (ride_length_temp_mins)

```

```{r}

ride_length_temp_mins_int = as.integer (ride_length_temp_mins)

```


```{r}
# Sanity check

table (ride_length_temp_mins_int)

```

```{r}

ride_length_mins_vector= ride_length_temp_hours_int*60 + ride_length_temp_mins_int

```

```{r}
summary (ride_length_mins_vector)
```

Now, if you feel satisfied with the result of this transformation, you can add 'ride_length_mins_vector' as a new column in your data frame


Check day_of_week for erroneous data:

```{r}
 table (bike_share_Oct22Feb23$day_of_week)
```

