---
title: "Bike Share Analysis"
output: html_notebook
---


```{r}

file_names_bike_share= dir("C:\\Users\\dsmit\\OneDrive\\Documents\\MDC Data Analytics\\Capstone\\Bike_Share_First.Rmd", full.names=TRUE)

```
```{r}
install.packages("tidyverse")
```


```{r}

library ("tidyverse")

```


```{r}
library("readxl")
```



```{r}

bike_share_raw_Oct22Feb23 <- do.call(rbind,lapply(file_names_bike_share,read_excel))

```


Take a look at the data frame:


```{r}

str(bike_share_raw_Oct22Feb23)

```

### Are there missing values?

```{r}
sum (complete.cases(bike_share_raw_Oct22Feb23))
```

Get a count of missing values per column (alternative 1):

```{r}

sapply(bike_share_raw_Oct22Feb23, function(x) {sum(is.na(x))})

```


Get a count of missing values per column (alternative 2):

```{r}

bike_share_raw_Oct22Feb23 %>% select(everything()) %>%summarise_all(funs(sum(is.na(.))))

```


How to handle the missing values?

a) Can we find a way of finding and entering the missing values?
b) Remove rows and find new rows if sample size is a concerned
c) Remove rows and use the reduced data frame (THIS WILL BE OUR APPROACH)

Other methods to handle missing values are:
Mean imputation (discuss)
Model imputation (discuss)
Regression imputation (discuss)

Are any of these last three methods applicable here?


```{r}

bike_share_Oct22Feb23 <- bike_share_raw_Oct22Feb23[complete.cases(bike_share_raw_Oct22Feb23), ]

```


Checking that the new data frame has no missing values:

```{r}
nrow(bike_share_Oct22Feb23)
sum (complete.cases(bike_share_Oct22Feb23))

```


### Check for incorrect values


```{r}
str(bike_share_Oct22Feb23)
```

Any strange values in the qualitative variables?

Check rideable_type and member_casual.

ASK STUDENTS: What am I trying to accomplish with the following code?

```{r}
 table (bike_share_Oct22Feb23$rideable_type)
 table (bike_share_Oct22Feb23$member_casual)


```

I do not think we need to worry about strange values in ride_id. What do you think?

Also, because we are not going to use started_at and ended_at, we can ignore checking for strange values in them.


Check ride_length:

```{r}
summary(bike_share_Oct22Feb23$ride_length)
```

Something is wrong, the values in ride_lenght were supposed to be in the format hh:mm:ss, but we are getting a date value as well.


```{r}

head (bike_share_Oct22Feb23$ride_length, 10)

```

```{r}

ride_length_temp= separate (bike_share_Oct22Feb23, ride_length, into= c("date_ride", "time_ride"), sep=" ")$time_ride

```

```{r}
head (ride_length_temp, 10)
```

Now, grab the hours and minutes (ignore seconds???) and convert them to an integer


