---
title: "Data frames"
output: html_notebook
---

# Data frames

A data frame is a matrix-like object with columns that can have different data types.

A data frame is different from a matrix (an R data structure that we will not cover). Matrices, like vectors, only accept columns with the same data type. This is a limitation because, what if we want the person’s name in column 1 and the person’s salary in column 2? A matrix cannot handle this, but a data frame can.

__Example__: Create a data frame with the presidents’ height data. Download the file "Data Winners&Opponents" from Blackboard.

```{r}
# Just copy the data in the file and paste it here. Then, run this code chunk to create the vectors

presid_name= c("Obama","Bush","Bush","Clinton","Clinton","Bush Father","Reagan","Reagan","Carter","Nixon","Nixon","Johnson","Kennedy","Eisenhower","Eisenhower","Truman")

winner = c(185, 182, 182, 188, 188, 188, 185, 185, 177, 182, 182, 193, 183, 179, 179, 175)

opponent = c(175, 193, 185, 187, 188, 173, 180, 177, 183, 185, 180, 180, 182, 178, 178, 173)


```

Let's say we want to create another vector to store the election year. The year starts in 2008 and goes all the way back until 1948.

Use the next code chunk to create a vector called year with the year values that satisfy this problem.

```{r}

year= seq (from= 2008, to= 1948, by=-4)

```

Create a vector called “isWinnerTaller”, which is True when Winner > Opponent and False otherwise. Then, use “isWinnerTaller” to count how many times the winner has been taller than the opponent.

```{r}

isWinnerTaller= winner > opponent

isWinnerTaller

```

```{r}
# Count how many times the winner has been taller than the opponent.

sum(isWinnerTaller)

```

Use all the previous vectors to create a data frame with all the data. We can use the data.frame() function to create R data frames.

```{r}

df_presidents= data.frame(presid_name, year, winner, opponent, isWinnerTaller)
df_presidents

```

Let's learn some handy functions to work with data frames

```{r}
head(df_presidents)

```

```{r}
str (df_presidents)

```

```{r}
colnames(df_presidents)

```

```{r}
ncol(df_presidents)
nrow(df_presidents)
```

Let's learn how to add a new column to an existing data frame.

For example, add a column named _difference_ as the last column. This column stores the difference in height between winners and opponents.

```{r}
df_presidents$difference = winner - opponent

```

```{r}
df_presidents
```

Now, let's learn how to delete a column by using the index of the column in the data frame.

For example, delete the difference column by using its index.

```{r}
df_presidents [ , -6]


```

```{r}

df_presidents

```

The right way is to re-assign the deletion to the data frame

```{r}
df_presidents = df_presidents [, -6]
```

```{r}
df_presidents

```

A better way of deleting the last column of a data frame IF you want to do it more programmatically is:

df_presidents = df_presidents [, -ncol(df_presidents)]

DO NOT RUN THIS STATEMENT. I just want you to know that this is an alternative way of deleting the the last column of a data frame.


Let's add the column _difference_ again, so we can delete it one more time using a different method.

```{r}
df_presidents$difference = winner - opponent

```

```{r}

df_presidents
```


Before, we deleted the column difference by using its index. Now, let's delete by using its name:

```{r}

df_presidents[ , colnames(df_presidents) != 'difference' ]

```

```{r}

df_presidents= df_presidents[ , colnames(df_presidents) != 'difference' ]

```

```{r}

df_presidents

```

### Indexing a data frame

Get the second column

```{r}

df_presidents[ , 2]

# alternative 2: df_presidents [, 'year']

# alternative 3: df_presidents$year

```

Get the last column

```{r}
df_presidents[ , ncol(df_presidents)]

# alternative: df_presidents[ , 5]

```


Get the winner’s and the opponent’s heights, only for the first three rows in the data set. 

```{r}
df_presidents [c(1,2,3), c(3,4)]

# alternative: df_presidents [1:3 , c(3,4)]

```

### Some useful functions to work with data frames

The subset() function

Example: Use the subset() function to get the rows from the data frame where the Winner > Opponent

```{r}

subset (df_presidents, df_presidents$isWinnerTaller==TRUE)

# or, more succinct: subset (df_presidents, isWinnerTaller)

```

Example: Get the winners’ names only for cases when Winner > Opponent

```{r}
subset (df_presidents$presid_name, df_presidents$isWinnerTaller==TRUE)

```

Other useful functions are apply() and tapply(). To be able to practice these functions, lets' add a column with the party of the winner (i.e., the party of the president).

```{r}

party= c("Dem", "Rep", "Rep", "Dem", "Dem", "Rep", "Rep", "Rep", "Dem","Rep","Rep","Dem","Dem","Rep","Rep","Dem")
  
```

```{r}

df_presidents$presid_party= party
df_presidents

```


Use tapply() to compute the mean height for the presidents from each party

```{r}
tapply (df_presidents$winner, df_presidents$presid_party, mean)

```

Use tapply() to compute the max height for the presidents from each party

```{r}
tapply (df_presidents$winner, df_presidents$presid_party, max)

```

Use apply() to compute the mean height for the winners and opponents

```{r}
apply (df_presidents [, c("winner", "opponent")] ,2, mean)

# The 2 means that we want to apply the mean by columns

```

We can also use the colMeans () function

```{r}
colMeans (df_presidents [, c("winner", "opponent")])
```

Use apply() to compute the sd of the height for the winners and opponents

```{r}
apply (df_presidents [, c("winner", "opponent")] , 2, sd)

```
```{r}
sd(df_presidents [, "winner"])

```

```{r}
sd(df_presidents [, "opponent"])
```






