---
title: "Simple Linear Regression"
output: html_notebook
---

### Example

#### Let’s find an equation to predict medv based on lstat

To be able to use the objects included in an R library, you need to load the library. Let's load the MASS library:

```{r}

library (MASS)

```

After the MASS library is loaded, you can access all its built-in objects (e.g., data sets, functions). Today, we only want to access the Boston data set part of MASS. Let's take a look at the Boston data set:

```{r}

str(Boston)
head(Boston)

```
Let's define the variables we are working with:

Y: medv (median house value in the neighborhood) (IT IS GIVEN IN UNITS OF $1000) 

X: lstat (percent of households with low socioeconomic status in the neighborhood)

<br>

#### Check if the linear relationship is reasonable

```{r}

# Obtain the sample correlation coefficient between X and Y

cor (Boston$lstat, Boston$medv)

# alternative code

with (Boston, cor (lstat, medv))

```
```{r}

# Scatter plot of Y vs X

with(Boston, plot (lstat,medv))

# alternative code: plot (Boston$lstat, Boston$medv))

# If you want to include the least square line in the plot, do the following:

abline(lm(medv ~ lstat, data = Boston), col = "blue")

```

__Interpretation of r and scatterplot___

The value of r suggests a fairly strong negative linear relationship between medv and lstat.

We can check in the scatterplot that the relationship is indeed negative. 

The scatterplot shows us that the linear relationship is acceptable (it fits the data points quite well), especially, in the range from 5< lstat< 30. However, towards the two endings, the linear model does not fit the data very well. A closer look at the graph shows that this non-linearity is because there is some curvature in the data (maybe a negative exponential relationship, or.. a polynomial)

But for sure, it is reasonable to continue with the linear regression analysis. Even if a linear might not be the BEST, it might give us good results based on the r and scatterplot analysis.

<br>

#### Apply Simple Linear Regression

```{r}

# We apply regression by using the lm() function, where lm stands for linear model.
# It is customary to save the result of applying lm() in a user-defined object

medv_simple_out= lm(medv ~ lstat, data=Boston)

# Then, we call the summary() function on the defined object

summary (medv_simple_out)

```
```{r}
# If you want to get the equation coefficients only, then:

medv_simple_out$coefficients

```
__Is this equation good?__

__Statistical significance__

Test for B1:

Ho: B1 = 0 (the slope of the real equation is zero; thus, there is no linear relationship between X and Y)

Ha: B1 != 0 (the slope of the real equation is different from zero; thus, there is a linear relationship between X and Y)

The t test statistic for the b1 coefficient is -24.53 (huge = very far from zero). Therefore, naturally, the PV is very small (PV < 2e-16). This PV is less than alpha. Therefore, we can reject Ho and support Ha. 

The data give us evidence to conclude that there is a statistically significant relationship between medv (Y) and lstat (X). 

OPTIONAL: You could also (optionally !!!) analyze the ANOVA table and the PV related to the F statistic. To get the ANOVA table, you would use the command: anova (medv_simple_out)

You would get to the same conclusion as we did wit the t test for B1.


__Interpret the slope of the equation (this interpretation is only done after we already showed statistical significance)__

The equation we got was:

Average medv of the neighborhood = 34.55 - 0.95*(lstat of the neighborhood)

Interpretation the slope: When thepercent of households with low socioeconomic status in a neighborhood increases by one percent, we expect the median household income to decrease on average by 950 dollars (0.95 = 950 because the medv values are given in units of 1000 dollars)


__Practical significance__ 

R squared is 0.5441 (54.4 %)

If we use the linear equation that we obtained instead of the sample mean to make predictions of Y, we can reduce 54.4 % of the variability (error level) in the values of Y (medv values). Since we reduce a little more than 50% of the variability compared to using the sample mean, __using the equation is better than using the sample mean to make predictions__.

Note: Remember that in regression, the term variability (or error) refers to the difference between the predicted values of Y given by the equation and real values of Y.


The RSE is 6.216 (= $ 6, 216)

```{r}
# If you want to get RSE programmatically, you can do this:

summary (medv_simple_out)$sigma

```

On average, the values of medv deviate from the regression equation by $ 6, 216

Stated in another way, we can expect predictions of medv based on the equation to be off by around $ 6, 216.

```{r}
summary(Boston$medv)

```
```{r}

# Let's compare RSE with the values of medv
# Are we going to compare RSE with all the values of medv?
# No, that' unfeasible
# A good strategy is to compare RSE with the average of medv
# That's precisely what the coeff of variation do. Let's compute the coff of variation:
# Coeff var = RSE/ sample mean

6.216/ mean(Boston$medv)

```
Converted into a %, the coeff of variation is 27.6%

An error of  $ 6, 216 does not seem to be as low as we would desire IN THIS PROBLEM. We should attempt to get a smaller RSE.

Partial conclusions:

The equation is statistically significant. It is better than the sample mean to make predictions of medv (based on R sq). The RSE is not as good as we would like. In any case, we are going to assume that we made the decision to use the equation to make predictions of medv.

Let's use the equation to predict medv for three new neighborhoods with lstat of 4.5 %, 5%, and 7%.
Let's use the predict() function to do this.

```{r}

predict(medv_simple_out, data.frame(lstat=c(4.5, 5, 7)))

```
What if we wanted to predict the medv values for the values of lstat in the Boston dataset?

```{r}

predict (medv_simple_out)

```


### Checking the regression assumptions

We can do a plot of the error terms (residuals) VS the predicted values of Y (or equivalently, VS the values of X)

```{r}

# The function residuals() allows us to estimate the error terms

plot (predict (medv_simple_out), residuals(medv_simple_out))

abline(h=0)

```
In this case, we can clearly see that the error terms show a non-linear pattern (a parabola shape).
Therefore, the relationship between medv and lstat does not seem to be linear. A linear equation muight be acceptable, but it is definitvely not the best option (because there is non linear pattern going on). Assumption 1 is not totally satisfied.

We are not going to check for assumption 4, because when the linear relationship is suspect, this graph is not suitable to check for assumption 4.

Now let's check for assumption 3 (Does the error term follow a Normal distribution?)

For that, you are going to apply a test to check for normal distribution. The test is called the Shapiro test

Ho: The variable follows a Normal distribution
Ha: The variable does not follow a Normal distribution

```{r}

shapiro.test(residuals(medv_simple_out))

```
The p-value < 2.2e-16, less than alpha, therefore, the error terms does not follow a normal distribution. Assumption 3 is not satisfied either.

THE END OF SIMPLE LINEAR RELATIONSHIP !!!






















