---
title: "Simple Linear Regression"
output: html_notebook
---

### Example

#### Let’s find an equation to predict medv based on lstat

To be able to use the objects included in an R library, you need to load the library. Let's load the MASS library:

```{r}

library (MASS)

```

After the MASS library is loaded, you can access all its built-in objects (e.g., data sets, functions). Today, we only want to access the Boston data set part of MASS. Let's take a look at the Boston data set:

```{r}

str(Boston)
head(Boston)

```
Let's define the variables we are working with:

Y: medv (median house value in the neighborhood) (IT IS GIVEN IN UNITS OF $1000) 

X: lstat (percent of households with low socioeconomic status in the neighborhood)

<br>

#### Check if the linear relationship is reasonable

```{r}

# Obtain the sample correlation coefficient between X and Y

cor (Boston$lstat, Boston$medv)

# alternative code

with (Boston, cor (lstat, medv))

```
```{r}

# Scatter plot of Y vs X

with(Boston, plot (lstat,medv))

# alternative code: plot (Boston$lstat, Boston$medv))

# If you want to include the least square line in the plot, do the following:

abline(lm(medv ~ lstat, data = Boston), col = "blue")

```

__Interpretation of r and scatterplot___

The value of r suggests a fairly strong negative linear relationship between medv and lstat.

We can check in the scatterplot that the relationship is indeed negative. 

The scatterplot shows us that the linear relationship is acceptable (it fits the data points quite well), especially, in the range from 5< lstat< 30. However, towards the two endings, the linear model does not fit the data very well. A closer look at the graph shows that this non-linearity is because there is some curvature in the data (maybe a negative exponential relationship, or.. a polynomial)

But for sure, it is reasonable to continue with the linear regression analysis. Even if a linear might not be the BEST, it might give us good results based on the r and scatterplot analysis.

<br>

#### Apply Simple Linear Regression

```{r}

# We apply regression by using the lm() function, where lm stands for linear model.
# It is customary to save the result of applying lm() in a user-defined object

medv_simple_out= lm(medv ~ lstat, data=Boston)

# Then, we call the summary() function on the defined object

summary (medv_simple_out)

```
```{r}
# If you want to get the equation coefficients only, then:

medv_simple_out$coefficients

```
__Is this equation good?__

__Statistical significance__

Test for B1:

Ho: B1 = 0 (the slope of the real equation is zero; thus, there is no linear relationship between X and Y)

Ha: B1 != 0 (the slope of the real equation is different from zero; thus, there is a linear relationship between X and Y)

The t test statistic for the b1 coefficient is -24.53 (huge = very far from zero). Therefore, naturally, the PV is very small (PV < 2e-16). This PV is less than alpha. Therefore, we can reject Ho and support Ha. 

The data give us evidence to conclude that there is a statistically significant relationship between medv (Y) and lstat (X). 

OPTIONAL: You could also (optionally !!!) analyze the ANOVA table and the PV related to the F statistic. To get the ANOVA table, you would use the command: anova (medv_simple_out)

You would get to the same conclusion as we did wit the t test for B1.


__Interpret the slope of the equation (this interpretation is only done after we already showed statistical significance)__

The equation we got was:

Average medv of the neighborhood = 34.55 - 0.95*(lstat of the neighborhood)

Interpretation the slope: When thepercent of households with low socioeconomic status in a neighborhood increases by one percent, we expect the median household income to decrease on average by 950 dollars (0.95 = 950 because the medv values are given in units of 1000 dollars)


__Practical significance__ 

R squared is 0.5441 (54.4 %)

If we use the linear equation that we obtained instead of the sample mean to make predictions of Y, we can reduce 54.4 % of the variability (error level) in the values of Y (medv values). Since we reduce a little more than 50% of the variability compared to using the sample mean, using the equation is better than using the sample mean to make predictions.

On Monday, we continue by analyzing RSE ...






