---
title: "Statistics Intro"
output: html_notebook
---

# Statistics Introduction

Statistics is the science of analyzing, reviewing and conclude data.

Some basic statistical numbers include:

Mean, median and mode
Minimum and maximum value
Percentiles
Variance and Standard Devation
Covariance and Correlation
Probability distributions

The R language was developed by two statisticians. It has many built-in functions, in addition to libraries for the exact purpose of statistical analysis.

# Data Set

A data set is a collection of data, often presented in a table.

There is a popular built-in data set in R called "mtcars" (Motor Trend Car Road Tests), which is retrieved from the 1974 Motor Trend US Magazine.
```{r}
# Print the mtcars data set

mtcars
```

You can use the question mark (?) to get information about the mtcars data set:
```{r}
# Use the question mark to get information about the data set

?mtcars
```

Result:

mtcars {datasets}	R Documentation
Motor Trend Car Road Tests
Description
The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models).

Usage
mtcars
Format
A data frame with 32 observations on 11 (numeric) variables.

[, 1]	mpg	Miles/(US) gallon
[, 2]	cyl	Number of cylinders
[, 3]	disp	Displacement (cu.in.)
[, 4]	hp	Gross horsepower
[, 5]	drat	Rear axle ratio
[, 6]	wt	Weight (1000 lbs)
[, 7]	qsec	1/4 mile time
[, 8]	vs	Engine (0 = V-shaped, 1 = straight)
[, 9]	am	Transmission (0 = automatic, 1 = manual)
[,10]	gear	Number of forward gears
[,11]	carb	Number of carburetors
Note
Henderson and Velleman (1981) comment in a footnote to Table 1: 'Hocking [original transcriber]'s noncrucial coding of the Mazda's rotary engine as a straight six-cylinder engine and the Porsche's flat engine as a V engine, as well as the inclusion of the diesel Mercedes 240D, have been retained to enable direct comparisons to be made with previous analyses.'

Source
Henderson and Velleman (1981), Building multiple regression models interactively. Biometrics, 37, 391-411.


require(graphics)
pairs(mtcars, main = "mtcars data", gap = 1/4)
coplot(mpg ~ disp | as.factor(cyl), data = mtcars,
       panel = panel.smooth, rows = 1)
## possibly more meaningful, e.g., for summary() or bivariate plots:
mtcars2 <- within(mtcars, {
   vs <- factor(vs, labels = c("V", "S"))
   am <- factor(am, labels = c("automatic", "manual"))
   cyl  <- ordered(cyl)
   gear <- ordered(gear)
   carb <- ordered(carb)
})
summary(mtcars2)

# Get Information

Use the dim() function to find the dimensions of the data set, and the names() function to view the names of the variables:
```{r}
Data_Cars <- mtcars # create a variable of the mtcars data set for better organization

# Use dim() to find the dimension of the data set
dim(Data_Cars)

# Use names() to find the names of the variables from the data set
names(Data_Cars)
```

Use the rownames() function to get the name of each row in the first column, which is the name of each car:
```{r}
Data_Cars <- mtcars

rownames(Data_Cars)
```

From the examples above, we have found out that the data set has 32 observations (Mazda RX4, Mazda RX4 Wag, Datsun 710, etc) and 11 variables (mpg, cyl, disp, etc).

A variable is defined as something that can be measured or counted.

Here is a brief explanation of the variables from the mtcars data set:

Variable Name	Description
mpg	Miles/(US) Gallon
cyl	Number of cylinders
disp	Displacement
hp	Gross horsepower
drat	Rear axle ratio
wt	Weight (1000 lbs)
qsec	1/4 mile time
vs	Engine (0 = V-shaped, 1 = straight)
am	Transmission (0 = automatic, 1 = manual)
gear	Number of forward gears
carb	Number of carburetors

# Print Variable Values

If you want to print all values that belong to a variable, access the data frame by using the $ sign, and the name of the variable (for example cyl (cylinders)):
```{r}
Data_Cars <- mtcars

Data_Cars$cyl
```

# Sort Variable Values

To sort the values, use the sort() function:
```{r}
Data_Cars <- mtcars

sort(Data_Cars$cyl)
```

From the examples above, we see that most cars have 4 and 8 cylinders.

# Analyzing the Data

Now that we have some information about the data set, we can start to analyze it with some statistical numbers.

For example, we can use the summary() function to get a statistical summary of the data:

```{r}
Data_Cars <- mtcars

summary(Data_Cars)
```

Do not worry if you do not understand the output numbers. You will master them shortly.

The summary() function returns six statistical numbers for each variable:

Min
First quantile (percentile)
Median
Mean
Third quantile (percentile)
Max
We will cover all of them, along with other statistical numbers in the next chapters.

# Max Min

The min() and max() functions can be used to find the lowest or highest value in a set:

Find the largest and smallest value of the variable hp (horsepower).
```{r}
Data_Cars <- mtcars

max(Data_Cars$hp)
min(Data_Cars$hp)
```

Now we know that the largest horsepower value in the set is 335, and the lowest 52.

We could take a look at the data set and try to find out which car these two values belongs to:

Observation of cars
                     mpg cyl  disp  hp drat    wt  qsec vs am gear carb
Mazda RX4           21.0   6 160.0 110 3.90 2.620 16.46  0  1    4    4
Mazda RX4 Wag       21.0   6 160.0 110 3.90 2.875 17.02  0  1    4    4
Datsun 710          22.8   4 108.0  93 3.85 2.320 18.61  1  1    4    1
Hornet 4 Drive      21.4   6 258.0 110 3.08 3.215 19.44  1  0    3    1
Hornet Sportabout   18.7   8 360.0 175 3.15 3.440 17.02  0  0    3    2
Valiant             18.1   6 225.0 105 2.76 3.460 20.22  1  0    3    1
Duster 360          14.3   8 360.0 245 3.21 3.570 15.84  0  0    3    4
Merc 240D           24.4   4 146.7  62 3.69 3.190 20.00  1  0    4    2
Merc 230            22.8   4 140.8  95 3.92 3.150 22.90  1  0    4    2
Merc 280            19.2   6 167.6 123 3.92 3.440 18.30  1  0    4    4
Merc 280C           17.8   6 167.6 123 3.92 3.440 18.90  1  0    4    4
Merc 450SE          16.4   8 275.8 180 3.07 4.070 17.40  0  0    3    3
Merc 450SL          17.3   8 275.8 180 3.07 3.730 17.60  0  0    3    3
Merc 450SLC         15.2   8 275.8 180 3.07 3.780 18.00  0  0    3    3
Cadillac Fleetwood  10.4   8 472.0 205 2.93 5.250 17.98  0  0    3    4
Lincoln Continental 10.4   8 460.0 215 3.00 5.424 17.82  0  0    3    4
Chrysler Imperial   14.7   8 440.0 230 3.23 5.345 17.42  0  0    3    4
Fiat 128            32.4   4  78.7  66 4.08 2.200 19.47  1  1    4    1
Honda Civic         30.4   4  75.7  52 4.93 1.615 18.52  1  1    4    2
Toyota Corolla      33.9   4  71.1  65 4.22 1.835 19.90  1  1    4    1
Toyota Corona       21.5   4 120.1  97 3.70 2.465 20.01  1  0    3    1
Dodge Challenger    15.5   8 318.0 150 2.76 3.520 16.87  0  0    3    2
AMC Javelin         15.2   8 304.0 150 3.15 3.435 17.30  0  0    3    2
Camaro Z28          13.3   8 350.0 245 3.73 3.840 15.41  0  0    3    4
Pontiac Firebird    19.2   8 400.0 175 3.08 3.845 17.05  0  0    3    2
Fiat X1-9           27.3   4  79.0  66 4.08 1.935 18.90  1  1    4    1
Porsche 914-2       26.0   4 120.3  91 4.43 2.140 16.70  0  1    5    2
Lotus Europa        30.4   4  95.1 113 3.77 1.513 16.90  1  1    5    2
Ford Pantera L      15.8   8 351.0 264 4.22 3.170 14.50  0  1    5    4
Ferrari Dino        19.7   6 145.0 175 3.62 2.770 15.50  0  1    5    6
Maserati Bora       15.0   8 301.0 335 3.54 3.570 14.60  0  1    5    8
Volvo 142E          21.4   4 121.0 109 4.11 2.780 18.60  1  1    4    2

By observing the table, it looks like the largest hp value belongs to a Maserati Bora, and the lowest belongs to a Honda Civic.

However, it is much easier (and safer) to let R find out this for us.

For example, we can use the which.max() and which.min() functions to find the index position of the max and min value in the table:
```{r}
Data_Cars <- mtcars

which.max(Data_Cars$hp)
which.min(Data_Cars$hp)
```

Or even better, combine which.max() and which.min() with the rownames() function to get the name of the car with the largest and smallest horsepower:
```{r}
Data_Cars <- mtcars

rownames(Data_Cars)[which.max(Data_Cars$hp)]
rownames(Data_Cars)[which.min(Data_Cars$hp)]
```

Now we know for sure:

Maserati Bora is the car with the highest horsepower, and Honda Civic is the car with the lowest horsepower.

# Outliers

Max and min can also be used to detect outliers. An outlier is a data point that differs from rest of the observations.

Example of data points that could have been outliers in the mtcars data set:

If maximum of forward gears of a car was 11
If minimum of horsepower of a car was 0
If maximum weight of a car was 50 000 lbs

# Mean, Median, and Mode

In statistics, there are often three values that interests us:

Mean - The average value
Median - The middle value
Mode - The most common value

# 
Mean
To calculate the average value (mean) of a variable from the mtcars data set, find the sum of all values, and divide the sum by the number of values.

Sorted observation of wt (weight)
1.513	1.615	1.835	1.935	2.140	2.200	2.320	2.465
2.620	2.770	2.780	2.875	3.150	3.170	3.190	3.215
3.435	3.440	3.440	3.440	3.460	3.520	3.570	3.570
3.730	3.780	3.840	3.845	4.070	5.250	5.345	5.42

Luckily for us, the mean() function in R can do it for you:
```{r}
# Find the average weight (wt) of a car:
Data_Cars <- mtcars

mean(Data_Cars$wt)
```

# Median

The median value is the value in the middle, after you have sorted all the values.

If we take a look at the values of the wt variable (from the mtcars data set), we will see that there are two numbers in the middle:

Sorted observation of wt (weight)
1.513	1.615	1.835	1.935	2.140	2.200	2.320	2.465
2.620	2.770	2.780	2.875	3.150	3.170	3.190	3.215
3.435	3.440	3.440	3.440	3.460	3.520	3.570	3.570
3.730	3.780	3.840	3.845	4.070	5.250	5.345	5.424
Note: If there are two numbers in the middle, you must divide the sum of those numbers by two, to find the median.

Luckily, R has a function that does all of that for you: Just use the median() function to find the middle value:
```{r}
# Find the mid point value of weight (wt):

Data_Cars <- mtcars

median(Data_Cars$wt)
```
# Mode

The mode value is the value that appears the most number of times.

R does not have a function to calculate the mode. However, we can create our own function to find it.

If we take a look at the values of the wt variable (from the mtcars data set), we will see that the numbers 3.440 are often shown:

Sorted observation of wt (weight)
1.513	1.615	1.835	1.935	2.140	2.200	2.320	2.465
2.620	2.770	2.780	2.875	3.150	3.170	3.190	3.215
3.435	3.440	3.440	3.440	3.460	3.520	3.570	3.570
3.730	3.780	3.840	3.845	4.070	5.250	5.345	5.424

Instead of counting it ourselves, we can use the following code to find the mode:
```{r}
Data_Cars <- mtcars

names(sort(-table(Data_Cars$wt)))[1]
```
# Percentiles

Percentiles are used in statistics to give you a number that describes the value that a given percent of the values are lower than.

If we take a look at the values of the wt (weight) variable from the mtcars data set:

Observation of wt (weight)
1.513	1.615	1.835	1.935	2.140	2.200	2.320	2.465
2.620	2.770	2.780	2.875	3.150	3.170	3.190	3.215
3.435	3.440	3.440	3.440	3.460	3.520	3.570	3.570
3.730	3.780	3.840	3.845	4.070	5.250	5.345	5.424

What is the 75. percentile of the weight of the cars? The answer is 3.61 or 3 610 lbs, meaning that 75% or the cars weight 3 610 lbs or less:
```{r}
Data_Cars <- mtcars

# c() specifies which percentile you want
quantile(Data_Cars$wt, c(0.75))
```

If you run the quantile() function without specifying the c() parameter, you will get the percentiles of 0, 25, 50, 75 and 100:
```{r}
Data_Cars <- mtcars

quantile(Data_Cars$wt)
```

# Quartiles

Quartiles are data divided into four parts, when sorted in an ascending order:

The value of the first quartile cuts off the first 25% of the data
The value of the second quartile cuts off the first 50% of the data
The value of the third quartile cuts off the first 75% of the data
The value of the fourth quartile cuts off the 100% of the data
Use the quantile() function to get the quartiles.