---
title: "Comparison of Two Means. Exercises"
output: html_notebook
---

__Note:__ To get the data for the two practice exercises, do the following:

1) Download the files “Two_Means_Ex1.rda” and “Two_Means_Ex2.rda” from Blackboard and do one of the following:

Option1: If you created an R project as I suggested at the beginning of the semester, move these two files __to the folder where this project is.__

Option 2: If you did not create a project, then move these two files to your working directory.
 

2) Load the files in R by running the following two statements in the Console:

load("Two_Means_Ex1.rda")

load("Two_Means_Ex2.rda")

After loading the files, the following vectors will be available in your R session:

domestic_rate, foreign_rate, coffeA, coffeeB


__Exercise 1)__ A study examined whether the perception of the quality of service at a five-star hotel in Jamaica differed by the type of guest (domestic vs foreign). Hotel guests were randomly selected and asked to rate 10 service-related items. Each item was rated on a five-point scale (the higher the score, the higher the perceived quality of service). A score for each guest was obtained by adding the ratings s/he gave for all items. The scores for domestic and foreign guests are recorded in the vectors “domestic_rate” and “foreign_rate”, which you already read into R.

Test whether the mean service rating given by domestic guests is higher than the mean service rating given by foreign guests using a 0.10 level of significance. Conduct this test step by step answering the following questions:

a) Define the two variables involved in this problem. 

Answer:

X1:

X2:

b) Check whether the assumptions needed for the test to be valid are satisfied.

Answer:


c) State the hypotheses (Ho and Ha)

Mx1:
Mx2:

Ho:
Ha: 

d) Conduct the relevant test in R to reach a decision. Assume the standard deviations of the two variables defined in part a) are the same.

DO INDIVIDUALLY. WORK ON IT INDIVIDUALLY FOR THREE MINS!

Hint: One of the arguments you have to pass to the function you need to use is: alternative = c("greater")



e) Compute the test statistic and P value for this problem without using the t.test () function. Use any other R built-in function but the t.test() function.

```{r}

# Here is the formula to get the Sp:

sp= sqrt ( ( 42*var(domestic_rate)+40*var(foreign_rate) ) / (43+41-2) )


```


f) What kind of error could you be making after you made the decision in part d)?


g) Estimate the effect size for this problem. In other words, estimate the magnitude of the effect that domestic guests have on the average rating when compared to foreign guests.

Note: DO PART G LATER, AFTER WE HAVE COVERED WHAT THE TERM "EFFECT SIZE" MEANS.


__Exercise 2)__ A Starbucks manager suspects that the coffee at two of her stores (stores A and B) yield different caffeine content. To test whether this is true, she takes a random sample of large coffees from both stores and measures the amount of caffeine content in each coffee cup (in mg). The amount of mg in a cup of coffee is a random variable that follows a Normal distribution. The results are stored in the variables “coffeeA” and coffeeB”. Conduct a hypothesis test to check if the manager’s concern is reasonable. Use a significance level of 0.05. Conduct the test following these steps:

- Define the two variables involved in this problem

A: 
B: 


- State the hypotheses (Ho and Ha)


- Use t.test () function to conduct the test


- State your decision 



b) What kind of error could you be making after you made that decision?



b1) What is the probability of committing such an error IF in reality the difference in the average caffeine content in the coffee from the two stores is 1 mg.


```{r}

#Share with students. Useful for b1):

spcoffe= sqrt ( ( 19*var(coffeeA)+ 19*var(coffeeB) ) / (38) )


```



