---
title: "Introduction to R vectors"
output: html_notebook
---

# Introduction

A vector is a type of object in R that allows us to store data in it. __All the data stored in a vector must have the same type__. Basic data types are integer, numeric, logical, and character.

__How to create a vector?__

a) Using the combine function: c()

```{r}
a= c(1,2,3,4,5)
a
a1= c(10,20,30)
a1

plot(a1)

```

You can use the combine function to create a new vector by combining multiple existing vectors

```{r}

a2= c(a, a1, 1000)

a2

```

b) Using the : function

```{r}

b= 10:20
b

d= 5:1
d

```

c) Using the seq () function

```{r}
e= seq(from=1, to=5)
e

```

Using the “by” argument with seq()

```{r}
f= seq(from=0, to=10, by=2)
f

```

What if you want the sequence 10, 8, 6,..., 0?

```{r}
f= seq(from=10, to=0, by= -2)
f

```

d) Using the rep () function

Let’s say you want to create a vector that contains the vector "a" (which we created earlier) repeated three times. 

```{r}
# One way of creating this vector is:

h= c(a,a,a)
h

```

```{r}
# A better way is to use the function rep()

h= rep (x=a, times=3)
h

```

```{r}
rep (x=c(a,a1), times=3)

```

Let’s say I want a new vector k to contain f repeated three times followed by the numbers 10 and 11. How to get it? Think for a minute about it

```{r}

k= c(rep(x=f, times=3), 10, 11)

k

```

Let’s use of the argument “each” in the function rep ()

```{r}
a
h1= rep (x=a, each=3)
h1

```

Using the arguments “times” and “each” together

```{r}
a
h2= rep (x=a, each=6, times= 3)
h2

```

e) Using the scan () function

Example: Create a vector called vectorx that contains the numbers 34, 56, and 66. Use the scan () function to create it.

```{r}
vectorx = scan()
vectorx

```

__Some Properties of a Vector__

The number of elements of a vector: length()
The data type of a vector: mode()

```{r}
a
length(a)
mode (a)

```

What happens if we added a character to the vector a?

```{r}

a= c(a,"w")
a
mode (a)

```

```{r}
b
b=c(b,"t")
b *2

```

What happened with the vector “a” was called “coercion”. Coercion is the process whereby R makes sure all the elements of a vector have the same datatype. Some common coercion examples:

* If you add a character to a numeric vector, all elements of the vector become character.

* If you add numbers to a vector with True or False values (i.e., a vector which type if Logical), the True or False are converted to numeric  (False = 0 and True =1); thus, the vector's data type changes from logical to numeric.


__Indexing a vector__

Indexing refers to the process of retrieving the elements at the different positions of a vector

Example: Let us create a new vector, with length 6. With positive and negative numbers. Call it vector_y

```{r}

vector_y= c(23, 5, 60, 4, -2, -9)

```


Get first element of vector_y

```{r}
vector_y[1]
```

Get the last element 

```{r}

vector_y[length(vector_y)]

```

Get first 3 elements 

```{r}
# Two possible alternatives
vector_y [c(1,2,3)]

vector_y [1:3]

```

Get the first and fifth elements

```{r}
vector_y [c(1,5)]
```

Get all but the second element of vector_y (long way)

```{r}

vector_y [ c(1, 3:length(vector_y)) ]   

# 3:length(vector_y... gives from position 3 and on

```

Easier way to retrieve all but the second element

```{r}
vector_y [-2]

```

Get all but the second and fifth elements

```{r}

vector_y [c(-2,-5)]

```


You can use one vector to index another vector

```{r}

i= c(3,4)
vector_y [i]

```

__Vectorized operations__

Many operations executed on vectors are vectorized, which means that a function applied to a vector is actually applied individually to each element of that vector.

Many (not all) R built-in functions are vectorized. For example, sqrt (), +, -, *, / … and others. Some functions are not vectorized since they return a single summary value, for example, sum () and min ().

```{r}

vector_y
vector_y + 2

```

```{r}

vector_y > 1

```

Two handy functions to apply on vectors:  all (), any ()

```{r}

any (vector_y > 1)

```

```{r}

all (vector_y > 1)

all (vector_y < 100)

```

__Vector Recycling__

Recycling: When you are applying an operation to two vectors, if this operation requires them to be the same length, R automatically recycles, or repeats, the shorter one, until it is long enough to match the longer one.

```{r}

c(1,2,3) + c(6,0,9,20,22)

# In the background, R does this: c(1,2,3,1,2) + c(6,0,9,20,22)

```

## Exercise: Practice with vectors

We are going to generate a sample of 10 values. These 10 values are going to be 10 random numbers from 1 to 50.

a) We are going to use the R built-in function sample() to generate these values
b) We are going to save these 10 values in a vector called random_sample
c) We are going to compute the mean of these values using the mean() function
d) We are going to compute the standard deviation of these values using the sd() function
e) We are going to compute the mean of these values WITHOUT using the mean() function
f) We are going to compute the standard deviation of these values WITHOUT using the sd() function

a) We are going to use the R built-in function sample() to generate these values

```{r}
# The next two lines allow us to generate 10 random values taken from the interval from 1 to 50

set.seed (1)
sample(x=1:50, size= 10)

```

b) We are going to save these 10 values in a vector called random_sample

```{r}

set.seed (1)
random_sample= sample(x=1:50, size= 10)
random_sample

```

c) We are going to compute the mean (average) of these values using the mean() function

```{r}

mean(random_sample)

```

d) We are going to compute the standard deviation of these values using the sd() function

```{r}

sd (random_sample)

```

e) We are going to compute the mean (average) of these values WITHOUT using the mean() function

```{r}

sum(random_sample) / length(random_sample)

# Mean= Sum of all the values / number of values

```

f) We are going to compute the standard deviation of these values WITHOUT using the sd() function

```{r}
topSD= sum( (random_sample - mean(random_sample)) ^ 2 ) 
bottomSD= length(random_sample) - 1
sqrt(topSD / bottomSD)

```














