---
title: "RANSOMWARE_RANSOMBO"
output: html_notebook
---

Please be aware that the comments below are for those groups in which the response variable has only one category, Ransomware.

At this point, most teams are working with datasets whose response variables have two different categories e.g (benign and defacement).


Every output obtained is for the label **RANSOMWARE_RANSOMBO**. Therefore any potential research question should start with the inclusion of the phrase **Given that the label is RANSOMWARE_RANSOMBO we predict (or we think) that.....**. I would like to mention that the two clusters obtained at the end of this tutorial will need further analysis and/or correction and/or customization.

We first read the file and explore the number of columns present in the dataset

```{r}
final_2<-read.csv("group5.csv",sep = ",",header = TRUE)
print(final_2)
```

I now decided to work certain group of columns. You can make a totally different decision.

```{r}
final_22<-final_2[,c(2,3,4,5,6,8,9,10,11,12)]
```

I now want to inspect the new data set

```{r}
print(final_22)
```


Let us run the function summary in order to get more details about our partial dataset.

```{r}
summary(final_22)
```

```{r}
str(final_22)
```

Based on the result of the lines of code ran above I nwo want to explore two variables, Destination.Port and Protocol.

```{r}
print(final_22$Destination.Port)
```

The mode for the field Destination.Port seems to be 443. Therefore, let us create a categorical variable because R is reading this fied as continuous.

Let us now apply some data manipulation techniques in order to have two classes for Destination.Port and Protocol. As I mentioned above, you can take  a totally different path.

```{r}
# Check if there are any rows with Destination.Port == "443"
if (any(final_22$Destination.Port == "443")) {
  final_22$Destination.Port <- ifelse(final_22$Destination.Port == "443", 1, 10708)
} else {
  # Handle the case when there are no matches
  final_22$Destination.Port <- NA  # or any other default value
}
```


```{r}
# final_22$Destination.Port <- ifelse(final_22$Destination.Port == "443", 1, 10708)
```



```{r}
final_22$Destination.Port <- factor(final_22$Destination.Port)
```

Let us confirm that the tasks just completed above ended in success by running the View function.

```{r}
View(final_22$Destination.Port)
```


```{r}
# final_22$Protocol <- ifelse(final_22$Protocol == "6", 1, 0)
# Check if there are any non-NA elements with the value "6" in the Destination.Port column
if (any(final_22$Destination.Port == "6", na.rm = TRUE)) {
  # Replace "6" with 1, and all other non-NA values with 10708
  final_22$Destination.Port <- ifelse(final_22$Destination.Port == "6", 1, 10708)
} else {
  # Handle the case when there are no matches or all values are NA
  final_22$Destination.Port <- NA  # or any other default value
}

```

```{r}
final_22$Protocol <- factor(final_22$Protocol)
```


Let us confirm the output is the one we are looking for the variable Protocol.

```{r}
View(final_22$Protocol)
```

Let us install (if necessary)the packages needed for this activity.

```{r}
pkg <- c("ggplot2", "RColorBrewer")
new.pkg <- pkg[!(pkg %in% installed.packages())]
if (length(new.pkg)) {
  install.packages(new.pkg)  
}
```


Let us use ggplot to analyze the Packets flow as a function of the Destination.Port knowing the label is **RANSOMWARE_RANSOMBO**.

```{r}
library(ggplot2)
gg <- ggplot(final_22, aes(final_22$Total.Fwd.Packets, final_22$Total.Backward.Packets, color=Destination.Port))
gg <- gg + scale_color_brewer(palette="Set2")
gg <- gg + geom_point(size=3) + theme_bw()
print(gg)
```

You would need to add a title, to properly label the axis and also include a regression line if possible. In invite you to make comments and discuss the output in detail.



```{r}
library(ggplot2)
gg <- ggplot(final_22, aes(final_22$Total.Fwd.Packets, final_22$Total.Backward.Packets, color=Protocol))
gg <- gg + scale_color_brewer(palette="Set2")
gg <- gg + geom_point(size=3) + theme_bw()
print(gg)
```


You would need to add a title, to properly label the axis and also include a regression line if possible. It is important to mention that I am expecting you to include as much information as possible regarding the information security fields used for this project.

Before starting the clusters analysis let us set the number of data points to a comfortable range. In other words, we do not find the default conditions for this particular problem sufficient.

```{r}
options(max.print=1000000)
```

We are using fields 7 and 8 from our dataset and we are clustering by Destination.Port (two levels). Please make as many comments as possible about this.

```{r}
set.seed(20)
project_cluster <- kmeans(final_22[, 7:8], 2, nstart = 20)
project_cluster
```


Make comments about the output obtained above. Find as much documentation as possible and/or consult the instructor (myself) in order to support your answers.

```{r}
table(project_cluster$cluster, final_22$Destination.Port)
```


```{r}
project_cluster$cluster <- as.factor(project_cluster$cluster)
ggplot(final_22, aes(final_22$Total.Fwd.Packets, final_22$Total.Backward.Packets, color = project_cluster$cluster)) + geom_point()+ggtitle(paste( "clusters"))
```



I encourage you to add a a reasonable title, properly label the axis and include as much information as possible about the output.

 I remind you this is just some general approach to the problem. I encourage you to improve the accuracy of every tool we have used in this unsupervised machine learning example.
 
I am again exploring the two classes we used to support our decision.

```{r}
summary(final_22$Destination.Port)
```

```{r}
summary(final_22$Protocol)
```

Let us now cluster by Protocol.

```{r}
table(project_cluster$cluster, final_22$Protocol)
```


```{r}
project_cluster$cluster <- as.factor(project_cluster$cluster)
ggplot(final_22, aes(final_22$Total.Fwd.Packets, final_22$Total.Backward.Packets, color = project_cluster$cluster)) + geom_point()+
 ggtitle(paste( "clusters"))
  

```



In order to make corrections to the output obtained in the cluster above we can redefine both  Protocol and Destination Port in order to obtain a reasonble results. Otherwise, we could subset our dataset in order to obtain a more accurate output. The idea of this tutorial is to get you started and/or contribute to the work you have done already.



