---
title: "Handout10_Ch5_Vis_Reg aka handout9"
output: html_notebook
---

--Welcome to Lecture#8,Chapter V.

Let us start by finding our working directoty.

```{r}
getwd()
```
# here i get to my directory

```{r}
# make sure the packages for this chapter
# are installed, install if necessary
pkg <- c("ggplot2", "scales", "maptools",
              "sp", "maps", "grid", "car" )
new.pkg <- pkg[!(pkg %in% installed.packages())]
if (length(new.pkg)) {
  install.packages(new.pkg)  
}
```

Let us now make sure that the file zeroaccess is uploaded into your new project. Then let us define the variable za.

# above i downloaded the necessary packages
```{r}
# read the CSV with headers
za <- read.csv("zeroaccess.csv", header=T,sep ="," )
#View(za)
View(za)
```

What is the output we have just obtained after running the chunk above.

# above got to read the csv files with headers and assigned to za
#View() allowed to see the file in another tab in RStudio

```{r}
# Load ggplot2 to create graphics
library(ggplot2)
# create a ggplot instance with zeroaccess data
gg <- ggplot(data=za, aes(x=long, y=lat)) 
# add the points, set transparency to 1/40th 
gg <- gg + geom_point(size=1, color="#000099", alpha=1/40) 
# add axes labels
gg <- gg + xlab("Longitude") + ylab("Latitude")
# simplify the theme for aesthetics
gg <- gg + theme_bw() 
# this may take a while, over 800,000 points plotted
print(gg)
```
# above is the load for graphic ggplot
# some css code for aesthetics
# some code for gps
# a map with plots and it did take a while to load 

```{r}
install.packages("mapproj")

```
```{r}
install.packages("maps")

```

```{r}
library(mapproj)
```


```{r}
# requires package : ggplot2
# requires object: za (5-1)
# the "maps" and "mapproj" packages are used by ggplot
# load map data of the world
world <- map_data("world")
#Remove Antarctica
world <- subset(world, world$region!="Antarctica")
# load world data into ggplot object
gg <- ggplot(data=world, aes(x=long, y=lat))
# trace along the lat/long coords by group (countries)
gg <- gg + geom_path(aes(group=group), colour="gray70")
# now project using the mercator projection
# try different projections with ?mapproject
gg <- gg + coord_map("mercator", xlim=c(-200, 200))
# load up the ZeroAccess points, overiding the default data set
gg <- gg + geom_point(data=za, aes(long, lat), 
                      colour="#000099", alpha=1/40, size=1)
# remove text, axes ticks, grid lines and do gray border on white
gg <- gg + theme(text=element_blank(), 
                 axis.ticks=element_blank(),
                 panel.grid=element_blank(),
                 panel.background=element_rect(color="gray50",
                                               fill="white"))
print(gg)

```
# here removed Antarctica from the region World map
# added some styling
# noticed a lot of blue in US Europe India

Interpret the output above. 

#Let us proceed to define the file county.dataset found on blackboard.


```{r}
# read up census data per county
county.data <- read.csv("countydataset.csv", header=T,sep = ",")

```

# here we have census data for counties

```{r}
View(county.data)
```

# a seperate tab was opened with a data spreadsheet, basically

```{r}
set.seed(1)
# generate 200 random numbers around 10
input <- rnorm(200, mean=10)
summary(input)
```
?set.seed

Random Number Generation
Description
.Random.seed is an integer vector, containing the random number generator (RNG) state for random number generation in R. It can be saved and restored, but should not be altered by the user.

RNGkind is a more friendly interface to query or set the kind of RNG in use.

RNGversion can be used to set the random generators as they were in an earlier R version (for reproducibility).

?rnorm

The Normal Distribution
Description
Density, distribution function, quantile function and random generation for the normal distribution with mean equal to mean and standard deviation equal to sd.

Interpret the syntax and the output above.

# set the seed to 1 
# assigned the normal distribution between 200 around 10 to input
# got a summary report for the input
# mean comes out as 10.036

```{r}
# requires objects: input (5-16)
# generate output around a mean of 2 x input
output <- rnorm(200, mean=input*2)
# put into data frame to plot it
our.data <- data.frame(input, output)
gg <- ggplot(our.data, aes(input, output))
gg <- gg + geom_point()
gg <- gg + geom_smooth(method = "lm", se=F, color="red")
gg <- gg + theme_bw()
print(gg)
```
# here use the input and output to plot a chart
# chart shows a positive linear relationship

Make comments about both the syntax and the output of the task executed above. Is it possible to customize this graph to make it more explanatory? If so explain and run the code.

On graphs, the explanatory variable is conventionally placed on the x-axis, while the response variable is placed on the y-axis.

To make the more explanatory I would need to add more predictors or more ways of cleaning up the x axis

Let us now run **output** vs **input**.

```{r}
model <- lm(output ~ input)
summary(model)
```
significance code         p-value
   ***                 [0, 0.001]
    **              (0.001, 0.01]
     *               (0.01, 0.05]
     .                (0.05, 0.1]
                         (0.1, 1] 
                         
Is the input significant at a 5% significance level?1%?

looks like the 1%

?lm
Fitting Linear Models
Description
lm is used to fit linear models, including multivariate ones. It can be used to carry out regression, single stratum analysis of variance and analysis of covariance (although aov may provide a more convenient interface for these)

Let us now proceed to get the confidence interval for this model.

```{r}
confint(model)
```
?confint

Confidence Intervals for Model Parameters
Description
Computes confidence intervals for one or more parameters in a fitted model. There is a default and a method for objects inheriting from class "lm".


Using the county.data dataset, let us run a model of Infections vs ufo2010(Aliens Visits according to the UFO).

```{r}
summary(lm(county.data$Infections ~ county.data$ufo2010, data= county.data))
```
Given the output printed above, is ufo2010 significant at a 5% significance level?1%?

# it has 3 stars so 1%


Let us now run a model of Infections vs every single quantitative variable that is included in the dataset.


```{r}
View(county.data)
```


```{r}
summary(lm(Infections ~ pop + income + ipaddr + ufo2010, 
           data=county.data))
```
# used more predictors 

Interpret the output. How would you proceed from now on in this handout given the results obtained above. 


```{r}
install.packages("carData")
```


```{r}
library(car) # for the vif() function
```

Let us just explore the variance inflation factor(VIF) of the model to see if there is a chance of high correlation between my predictors. I remind you that a strong correlation between two of my predictors will likely end up in heteroskedasticity, and therefore our model would not be accuarate.  

```{r}
model <- lm(Infections ~ pop + income + ipaddr + ufo2010, 
            data=county.data)
sqrt(vif(model))
```
?sqrt

Miscellaneous Mathematical Functions
Description
sqrt(x) computes the (principal) square root of x, \sqrt{x} 

The naming follows the standard for computer languages such as C or Fortran.

Let us see if population affects the number of infections in this dataset. Write the null and alternative hypothesis you would use to test this relationship. 


```{r}
summary(lm(Infections ~ pop, data=county.data))
```

Interpret the results obtained above. 


Now let us define the regression of Infections vs pop as pop.lm and predict the number of infections based on the variable population. 


```{r}
pop.lm <- lm(Infections ~ pop, data=county.data)
predict(pop.lm, data.frame(pop=6000000), interval="confidence")
```
?predict

Model Predictions
Description
predict is a generic function for predictions from the results of various model fitting functions. The function invokes particular methods which depend on the class of the first argument.

Interpret the results obtained above.

