---
title: "Handout_11"
output:
  html_document:
    df_print: paged
---

ctrl+alt+r
read all

```{r}
data<-read.csv("quiz2.csv",header = TRUE,sep = ",")
data
```

read the csv file quiz2 with the headers
by just putting the name of created variable data
it gives me a readout i think is better than View() 

```{r}
summary(data)
```

use summary() with data to get a read out 
subregion and region are chars
the rest are int

```{r}
str(data)
```

i get another read out which i like str()
i can see the data types associated to each var
29 obs of 7 vars

```{r}
head(data)
```

head() will give me a readout of the top rows of the dataset

```{r}
model1<-lm(infections~income+pop,data = data)
summary(model1)
```

get anova read out with income and pop as x
infections as the DV
assign to model1

```{r}
confint(model1)
```

i can use confint()
?confint

Computes confidence intervals for one or more parameters in a fitted model. There is a default and a method for objects inheriting from class "lm"

Let us proceed to remove regions and subregions and test for multicolinearility. 


```{r}
data1<-data[,c(3,4,5,6,7)]
data1
```

i removed the first 2 columns by not including them when i assigned the df to new var
data1
i like this way much better than Vew()

```{r}
cor.data<-cor(data1)
cor.data
```
cor() computes the correlation of x and y 
assign to cor.data
get the read out
the values closest to 1 have the strongest positive relationship

```{r}
#install.packages("Correlplot")

install.packages("contrib.url")
install.packages("corrplot")
```

```{r}
library(corrplot)
```

R package corrplot provides a visual exploratory tool on correlation matrix that supports automatic variable reordering to help detect hidden patterns among variables.

corrplot is very easy to use and provides a rich array of plotting options in visualization method, graphic layout, color, legend, text labels, etc. It also provides p-values and confidence intervals to help users determine the statistical significance of the correlations.

```{r}
corrplot(cor.data,method = "ellipse")
```

**Visualization methods in corrplot**

There are seven visualization methods (parameter method) in corrplot package, named "circle", "square", "ellipse", "number", "shade", "color", "pie".

Positive correlations are displayed in blue and negative correlations in red color. Color intensity and the size of the circle are proportional to the correlation coefficients.

Let us now proceed to install the leap package in order to do further analysis of the model. 

closest to one is the correlation between ip address and ufo 2010 sightings

```{r}
install.packages("leaps")
library("leaps")
```

install packages leaps

leaps() performs an exhaustive search for the best subsets of the variables in x for predicting y in
linear regression, using an efficient branch-and-bound algorithm. It is a compatibility wrapper for
regsubsets does the same thing better.
Since the algorithm returns a best model of each size, the results do not depend on a penalty model
for model size: it doesn’t make any difference whether you want to use AIC, BIC, CIC, DIC, ...```{r}
model2<-lm(infections~.,data = data1)
summary(model2)

Let us proceed to use regsubsets (a function for model selection)

```{r}
submodel2<-regsubsets(infections~.,data = data1)
best.summary<-summary(submodel2)
best.summary
```

?regsubsets
Model selection by exhaustive search, forward or backward stepwise, or sequential replacement

use the regsubsets to see which predictors will display correlations with * 
the more * the stronge the relationship

ufo2010 has the most *
ranked 1

```{r}
names(best.summary)
```

?names
Functions to get or set the names of an object.

use names() to get the best available functions to get or set names of a var

```{r}
which.min(best.summary$rss)
```

which.min() will give me the min value for the calculated residual sum of squares 
which is 4

```{r}
par(mfrow=c(1,2))
```

?par
par can be used to set or query graphical parameters. Parameters can be set by specifying them as arguments to par in tag = value form, or by passing them as a list of tagged values

?mfrow
The mfrow and mfcol parameters allow you to create a matrix of plots in one plotting space. Both parameters take a vector of length two as an argument, corresponding to the number of rows and columns in the resulting plotting matrix. 

```{r}
plot(best.summary$cp,xlab = "number of features",ylab = "cp")
```

plotting the chart with number of feature on the x axis
plotting the chart with cp on the y axis
using the best.summary variable

```{r}
which.min(best.summary$bic)
```

The Bayesian Information Criterion, often abbreviated BIC, is a metric that is used to compare the goodness of fit of different regression models

In practice, we fit several regression models to the same dataset and choose the model with the lowest BIC value as the model that best fits the data

adjusted best.summary with bic which is a metric that is used to compare the goodness of fit of different regression models 

use which.min() to find the min value = 1

```{r}
which.max(best.summary$adjr2)
```

adjR2: Adjusted coefficient of determination

Calculates the adjusted coefficient of determination of a multiple linear regression model

adjusted best.summary with adjr2 which calculates the adjusted coefficient of determination of a multiple linear regression model

use which(max) to find the max value =2

```{r}
bestsummary.fit1<-lm(infections~income+ufo2010,data = data1)
```

use lm() to use anova with infections as DV(y) and income(x) + ufo2010(x)

```{r}
summary(bestsummary.fit1)
```

Make a decision that you think can help us get the best results out of this model.

with the read out i see Ho the null hypothesis as being correct beacuse no p-value is lower than alpha

so i think i need to add more x(predictors) to the equation

