---
title: "Machine_Learning"
output: html_notebook
---
These packages are needed for this assignment:

```{r}
pkg <- c("ggplot2", "RColorBrewer")
new.pkg <- pkg[!(pkg %in% installed.packages())]
if (length(new.pkg)) {
  install.packages(new.pkg)  
}
```


Let us first make sure we are working in a directory we feel comfortable with. So I will start by changing my working directory to the one listed below


Let us now read the file and start exploring the dataset *memproc*.

```{r}
mydata1<- read.csv("memproc.csv", header=T)
summary(mydata1)
```

In order to explore this dataset more in detail, let us create a plot to compare the processor and memory usage, and differentiate it based on the malware state.

```{r}
library(ggplot2)
gg <- ggplot(mydata1, aes(proc, mem, color=state))
gg <- gg + scale_color_brewer(palette="Set2")
gg <- gg + geom_point(size=3) + theme_bw()
print(gg)
```


I sugest you to create now a new graph but this time including the title in the chart. The title would be "Memory vs Processor Usage as function of the Malaware state".


```{r}
set.seed(1492)
# count how many in the overall sample
n <- nrow(mydata1)
# set the test.size to be 1/3rd
test.size <- as.integer(n/3)
# randomly sample the rows for test set
testset <- sample(n, test.size)
# now split the data into test and train
test <- mydata1[testset, ]
train <- mydata1[-testset, ]
```



```{r}
# pull out proc and mem columns for infected then normal
# then use colMeans() to means of the columns
inf <- colMeans(train[train$state=="Infected", c("proc", "mem")])
nrm <- colMeans(train[train$state=="Normal", c("proc", "mem")])
```




```{r}
print(inf)
print(nrm)
```



```{r}
predict.malware <- function(data) {
  # get 'proc' and 'mem' as numeric values
  proc <- as.numeric(data[['proc']])
  mem <- as.numeric(data[['mem']])
  # set up infected comparison
  inf.a <- inf['proc'] - proc
  inf.b <- inf['mem'] - mem
  # pythagorean distance c = sqrt(a^2 + b^2)
  inf.dist <- sqrt(inf.a^2 + inf.b^2)
  # repeat for normal systems
  nrm.a <- nrm['proc'] - proc
  nrm.b <- nrm['mem'] - mem
  nrm.dist <- sqrt(nrm.a^2 + nrm.b^2)
  # assign a label of the closest (smallest)
  ifelse(inf.dist<nrm.dist,"Infected", "Normal")
}
# could test with these if you uncomment them
# predict.malware(inf['proc'], inf['mem'])
# expect "Infected" 
# predict.malware(nrm['proc'], nrm['mem'])
# expect "Normal"
```


```{r}
prediction <- apply(test, 1, predict.malware)
```



```{r}
sum(test$state==prediction)/nrow(test)
```



```{r}
# Figure 9-2 #########################################################
slope <- -1*(1/((inf['mem']-nrm['mem'])/(inf['proc']-nrm['proc'])))
intercept <- mean(c(inf['mem'], nrm['mem'])) - (slope*mean(c(inf['proc'], nrm['proc'])))
```



```{r}
result <- cbind(test, predict=prediction)
result$Accurate <- ifelse(result$state==result$predict, "Yes", "No")
result$Accurate <- factor(result$Accurate, levels=c("Yes", "No"), ordered=T)
```



```{r}
gg <- ggplot(result, aes(proc, mem, color=state, size=Accurate, shape=Accurate))
gg <- gg + scale_shape_manual(values=c(16, 8))
gg <- gg + scale_size_manual(values=c(3, 6))
gg <- gg + scale_color_brewer(palette="Set2")
gg <- gg + geom_point() + theme_bw()
gg <- gg + geom_abline(intercept = intercept, slope = slope, color="gray80")
print(gg)
```



```{r}
set.seed(1)
x <- runif(200, min=-10, max=10)
y <- 1.377*(x^3) + 0.92*(x^2) + .3*x + rnorm(200, sd=250) + 1572
x <- x + 10
smooth <- ggplot(data.frame(x,y), aes(x, y)) + geom_point() + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 3), size = 1, se=F) + 
  theme_bw()
print(smooth)
```



```{r}
memproc <- read.csv("memproc.csv", header=T)
memproc$infected <- ifelse(memproc$state=="Infected", 1, 0)
set.seed(1492)
n <- nrow(memproc)
test.size <- as.integer(n/3)
testset <- sample(n, test.size)
test <- memproc[testset, ]
train <- memproc[-testset, ]
```



```{r}
glm.out = glm(infected ~ proc + mem, data=test, family=binomial(logit))
summary(glm.out)
modelog <- predict.glm(glm.out, test, type="response")
gg <- ggplot(data.frame(x=modelog, y=ifelse(test$infected>0.5, "Yes", "No")), aes(x, y)) +
  geom_point(size=3, fill="steelblue", color="black", shape=4) + 
  ylab("Known Infected Host") +
  xlab("Estimated Probability of Infected Host") + theme_bw()
print(gg)
```



```{r}
set.seed(1) # repeatable
x <- c(rnorm(200), rnorm(400)+2, rnorm(400)-2)
y <- c(rnorm(200), rnorm(200)+2, rnorm(200)-2, rnorm(200)+2, rnorm(200)-2)
randata <- data.frame(x=x, y=y)
out <- list()
for(i in c(3,4,5,6)) {
  km <- kmeans(randata, i)
  centers <- data.frame(x=km$centers[ ,1], y=km$centers[ ,2], cluster=1)
  randata$cluster <- factor(km$cluster)
  gg <- ggplot(randata, aes(x, y, color=cluster)) + geom_point(size=2)
  gg <- gg + geom_point(data=centers, aes(x, y), shape=8, color="black", size=4)
  gg <- gg + scale_x_continuous(expand=c(0,0.1))
  gg <- gg + scale_y_continuous(expand=c(0,0.1))
  gg <- gg + ggtitle(paste("k-means with", i, "clusters"))
  gg <- gg + theme(panel.grid = element_blank(),
                   panel.background = element_rect(colour = "black", fill=NA),
                   axis.text = element_blank(),
                   axis.title = element_blank(),
                   legend.position = "none",
                   axis.ticks = element_blank())
  out[[i-2]] <- gg
}
print(out[[1]])
print(out[[2]])
print(out[[3]])
print(out[[4]])
```


