---
title: "Handout_Last"
output: html_notebook
---

Good afternoon all! First let us find out what our working directory is.

```{r}
getwd()
```

Today we are going to query the table quiz2 by using Transact-SQL inside the R chunks. First, we are going to need to install the package **sqldf**.

```{r}
install.packages("sqldf")
```


```{r}
library(sqldf)
```


```{r}
data1<-read.csv("quiz2.csv",header = TRUE,sep = ",")
```


```{r}
summary(data1)
```

Let us run a query that returns every single record from the table data1.

```{r}
sqldf("SELECT * FROM data1")
```

Now I would like you to create a query that returns every single record (row) for the fields region and population.

Let us now create a query that returns the first 10 records of the data1 table.

```{r}
sqldf('SELECT * FROM data1 LIMIT 10')
```


```{r}
sqldf("SELECT * FROM data1 ORDER BY infections DESC  LIMIT 5")
```

Interpret the output of the query ran above and run a similar query (limit 7) including the attributes **ipadddr** , **ufo2010** ,and **infections**.

```{r}
sqldf('SELECT region,infections FROM data1 WHERE infections > 1000 order by 2 desc')
```

Now I would like you to inspect (create a query) income,region and population when the condition is income>=50,000. Make comments.

```{r}
sqldf('SELECT * FROM data1 WHERE (income > 50000 AND infections > 2000) OR pop > 100000')
```

Interpret the query ran above and create a query that recreates a similar case scenario (one that is critical for the success of this mini-project).

Let us run a query that returns every region and population with a number of infections between 1000 and 7000. 

```{r}
sqldf('SELECT region,pop FROM data1 WHERE infections between  1000 and 7000')
```

Interpret the query below and list a hypothetical case scenario in which this query would be useful.

```{r}
sqldf('SELECT * FROM data1 WHERE subregion LIKE "A%"')
```

Last,let us find the average number of infections in our table.

```{r}
sqldf("SELECT AVG(infections) FROM data1")
```

Let us inspect the new table defined below data2.

```{r}
data2<-read.csv("finalexam.csv",header = TRUE,sep = ",")
```

Let us apply summary to data2 and examine the output.
```{r}
summary(data2)
```

Next, I would like you to complete the following steps:

1-Run a query that returns full details for every record from the west region.

2-Run a query that returns full details for every record from the east region where the number of infections is greater than 1500.

3-Run a few queries that can potentially help us to make the point that:
   a.Region affects the number of infections
   b.Population affects the number of infections
   c.Income affects the number of infections
   d.ufo2010 affects the number of infections
   e. The different interactions of region and income, and income and population affects the number of infections.
   
Explain every single query you include in this part of the handout.


