--
title: "Handout_10_reg"
output: html_notebook
---

ctrl + alt + r = run all
no errors

```{r}
getwd()
```
got my directory ready and available 

```{r}
# make sure the packages for this chapter
# are installed, install if necessary
pkg <- c("ggplot2", "scales", "maptools",
              "sp", "maps", "grid", "car" )
new.pkg <- pkg[!(pkg %in% installed.packages())]
if (length(new.pkg)) {
  install.packages(new.pkg)  
}
```

here i'm making sure the packages are installed for this handout

```{r}
# read the CSV with headers
regression1<-read.csv("incidents.csv", header=T,sep =",")
```

i need to read the incidents csv file into r with its headers
assign it to regression1            
      
```{r}
#View(regression1)
View(regression1)
```

i can use the View() to see a spreadsheet in r on another tab

```{r}
summary(regression1)
```

here i got a summary report of regression1
area, zone, population and incidents
area, zone and population are characters
incidents has a quantitative read out

```{r}
str(regression1)
```

str() will give me a str read out
16 obs of 4 vars
area, zone and population are characters
incidents has a quantitative read out

```{r}
regression1$population <- as.numeric(gsub(",","",regression1$population))
regression1$population
```

i got convert population into a num using as.numeric()


```{r}
str(regression1$population)
```

i confirm that population is changed into a num with str()

```{r}
regression2<-regression1[,-1]#new data frame with the deletion of column 1 
```

now i assign the df with the deletion of area with regression2


```{r}
head(regression2)
```

i can confirm that the column has been deleted with head()

```{r}
reg.fit1<-lm(regression1$incidents ~ regression1$population)
```

lm() will carry out analysis of chosen columns and fit a linear model
assign to reg.fit1

```{r}
summary(reg.fit1)
```

use the summary() for reg.fit1 and it calls the lm()
summary after using the lm is much more valuable 
Residuals:
   Min     1Q Median     3Q    Max 
-684.5 -363.5 -156.2  133.9 1164.7 

Coefficients:
                        Estimate Std. Error t value Pr(>|t|)  
(Intercept)            4.749e+02  2.018e+02   2.353   0.0337 *
regression1$population 8.462e-05  5.804e-05   1.458   0.1669  
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 534.9 on 14 degrees of freedom
Multiple R-squared:  0.1318,	Adjusted R-squared:  0.0698 
F-statistic: 2.126 on 1 and 14 DF,  p-value: 0.1669

from the read out i can see the intercepts p-value is less than alpha

```{r}
reg.fit2<-lm(incidents ~ zone+population, data = regression1)
```

lets add zone and population to the mix 
anova 2 way
assign to reg.fit2

```{r}
summary(reg.fit2)
```

after adding zone and population to se if their relationship with incidents 
i get a report and i see all the good plus zonewest which is interesting

zone west p-value is less than alpha

```{r}
regression1$zone <- ifelse(regression1$zone == "west", 1, 0)
```

?ifelse

Description
ifelse returns a value with the same shape as test which is filled with elements selected from either yes or no depending on whether the element of test is TRUE or FALSE.

Usage
ifelse(test, yes, no)

check to see if the zone is west if yes 1 if no 0

```{r}
View(regression1)
```

with View() i can see the zone column with 1's and 0's for yes and no for west

```{r}
str(regression1)
```

with str() i can see the zone as a num column with 1's and 0's for yes and no for west

```{r}
regression1$zone<-as.integer((regression1$zone),replace=TRUE) 
```



```{r}
interaction<-regression1$zone*regression1$population
```

mulitple the zone and population
assign to interaction

```{r}
reg.fit3<-lm(regression1$incidents~interaction+regression1$population+regression1$zone)
```

run analysis to fit a model
use the interaction between zone and population + population and zone against incidents
assign to reg.fit3

```{r}
summary(reg.fit3)
```

use summary to get the read out to see if the relationships

regression1$zone p-value is less than alpha

```{r}
reg.fit4<-lm(regression1$incidents~interaction)
```

run analysis to fit a model
use the interaction between zone and population  against incidents
assign to reg.fit4

```{r}
summary(reg.fit4)
```

use summary to get the read out to see if the relationships

both the intercept and interaction p-value is less than alpha
