---
title: "Handout_12_reg-ML"
output:
  html_document:
    df_print: paged
---

```{r}
getwd()
```




```{r}
data<-read.csv("quiz2.csv",header = TRUE,sep = ",")
data
```


```{r}
summary(data)
```


```{r}
str(data)
```

```{r}
head(data)
```


```{r}
model1<-lm(infections~income+pop,data = data)
summary(model1)
```


```{r}
vif(model1)
```


```{r}
confint(model1)
```



```{r}
data1<-data[,c(3,4,5,6,7)]
```


```{r}
cor.data<-cor(data1)
cor.data
```




```{r}
install.packages("Corrplot")
```

```{r}
library(corrplot)
```

```{r}
corrplot(cor.data,method = "ellipse")
```



```{r}
install.packages("leaps")
```


```{r}
library("leaps")
```


```{r}
model2<-lm(infections~.,data = data1)
summary(model2)
```


```{r}
vif(model2)
```


```{r}
submodel2<-regsubsets(infections~.,data = data1)
best.summary<-summary(submodel2)
```



```{r}
names(best.summary)
```



```{r}
which.min(best.summary$rss)
```

```{r}
par(mfrow=c(1,2))
```

```{r}
plot(best.summary$cp,xlab = "number of features",ylab = "cp")
```


```{r}
which.min(best.summary$bic)
```



```{r}
which.max(best.summary$adjr2)
```


```{r}
bestsummary.fit1<-lm(infections~income+ufo2010,data = data1)
```



```{r}
summary(bestsummary.fit1)
```

Make a decision that you think can help us get the best results out of this model.


```{r}
par(mfrow=c(2,2))
plot(bestsummary.fit1)
```


```{r}
library(car)
vif(bestsummary.fit1)
```



```{r}
bestsummary.fit2<-lm(infections~ufo2010,data = data1)
summary(bestsummary.fit2)
```


```{r}
bestsummary.fit3<-lm(infections~income,data = data1)
summary(bestsummary.fit3)
```


Let us run the Breuch-Pagan (BP)test in order to address the heteroskedasticity issue:


$H_0$:$sigma^2=0$
$H_a$:$Otherwise$

```{r}
install.packages("lmtest")
```


```{r}
library(lmtest)
```



```{r}
bptest(bestsummary.fit2)
```

P-value =0.8132 so we do not have enough evidence to reject the null(error variance=0)


```{r}
plot(bestsummary.fit2$fitted.values,data1$infections,xlab = "predicted",ylab = "actual",main = "Actual vs Predicted")
```



```{r}
data1["Actual"]=data$ufo2010
#create the vector actual
data1$Forecast=predict(bestsummary.fit2)
#populate forecast with the predicted values
```



```{r}
library(ggplot2)
```




```{r}
ggplot(data1,aes(x=Forecast,y=Actual))+geom_point()
geom_smooth(method = lm)
```

I would like you to add a title to the plot obtained above.

We need this package in order to calculate the Prediction Error Sum of Squares.

```{r}
install.packages("MPV")
```


```{r}
library(MPV)
```


Let us examine the prediction sum of squares(Press) statistic and selecting the model that has the lowest value.

```{r}

PRESS(model2)#every quantitative predictor
PRESS(bestsummary.fit1)#income+ufo2010
PRESS(bestsummary.fit3)#income
PRESS(bestsummary.fit2)#ufo
```


#Other linear considerations
Let us discuss first the inclusion of a quantitative variable and second the inclusion of an interaction term.


```{r}
finaldata<-read.csv("finalexam.csv",header = TRUE,sep = ",")
finaldata
```


```{r}
install.packages("ISLR")
```


```{r}
library(ISLR)
```



```{r}
str(data)
```


```{r}
qualitative.fit<-lm(infections~region+ufo2010,data = finaldata)
summary(qualitative.fit)
```




```{r}
incufo<-finaldata$ufo2010*finaldata$income
```


```{r}
interaction.fit<-lm(infections~region+incufo,data = finaldata)
summary(interaction.fit)
```

```{r}
PRESS(qualitative.fit)
PRESS(interaction.fit)
```

We would have to make a decision when it comes to the model we would like to use.

