---
title: "R Notebook"
output: html_notebook
---


In this handout we are going to focus in graphing information about malicious nodes. As security data scientist professionals,it is a good idea to get into the habit of visualizing data to pick up structures or patterns we might not see otherwise.

```{r}
anscombe[,c(1,5,2,6,3,7,4,8)]
```
?anscombe

Anscombe's Quartet of ‘Identical’ Simple Linear Regressions
Description
Four xx-yy datasets which have the same traditional statistical properties (mean, variance, correlation, regression line, etc.), yet are quite different.
```{r}
sapply(anscombe,mean)
```
?sapply
Apply a Function over a List or Vector
Description
sapply is a user-friendly version and wrapper of lapply by default returning a vector, matrix or, if simplify = "array", an array if appropriate, by applying simplify2array(). sapply(x, f, simplify = FALSE, USE.NAMES = FALSE) is the same as lapply(x, f).

```{r}
sapply(anscombe,sd)
```



```{r}
sapply(anscombe,var)
```




```{r}
for (i in 1:4) cat(cor(anscombe[,i], anscombe[,i+4]), "\n")
```

What do these values above represent?Explain.



```{r}
# Retrieve and read ZeuS blocklist data into R
zeusURL <- "https://zeustracker.abuse.ch/blocklist.php?download=ipblocklist"
zeusData <- "data/zeus.csv"
if (file.access(zeusData)) {
  # need to change download method for universal "https" compatibility
  download.file(zeusURL, zeusData, method="curl") 
}
# read in the ZeuS table; skip junk; no header; assign colnames
zeus <- read.table(zeusData, skip=5, header=FALSE, col.names=c("IP"))
```

Explain both the output and the syntax of the task completed above.


```{r}
# requires packages: igraph, plyr, RColorBrewer, colorspace
# requires object: set2 (4-4)
library(igraph)
library(plyr)
library(colorspace)
```





```{r}
# HELPER FUNCTION MENTIONED IN THE BOOK
# BUT NOT IN THE PRINTED LISTINGS

trim <- function (x) gsub("^\\s+|\\s+$", "", x)
```


```{r}
# HELPER FUNCTION MENTIONED IN THE BOOK
# BUT NOT IN THE PRINTED LISTINGS

BulkOrigin <- function(ip.list,host="v4.whois.cymru.com",port=43) {
 
  # setup query
  cmd <- "begin\nverbose\n" 
  ips <- paste(unlist(ip.list), collapse="\n")
  cmd <- sprintf("%s%s\nend\n",cmd,ips)
  
  # setup connection and post query 
  con <- socketConnection(host=host,port=port,blocking=TRUE,open="r+")  
  cat(cmd,file=con)
  response <- readLines(con)
  close(con)
  
  # trim header, split fields and convert results
  response <- response[2:length(response)]
  response <- laply(response,.fun=function(n) {
    sapply(strsplit(n,"|",fixed=TRUE),trim)
  })
  response <- adply(response,c(1))
  response <- subset(response, select = -c(X1) )
  names(response) = c("AS","IP","BGP.Prefix","CC",
                      "Registry","Allocated","AS.Name")
  
  return(response)
  
}
```


```{r}
# HELPER FUNCTION MENTIONED IN THE BOOK
# BUT NOT IN THE PRINTED LISTINGS

BulkPeer <- function(ip.list,host="v4-peer.whois.cymru.com",port=43) {
   
  # setup query
  cmd <- "begin\nverbose\n" 
  ips <- paste(unlist(ip.list), collapse="\n")
  cmd <- sprintf("%s%s\nend\n",cmd,ips)
  
  # setup connection and post query
  con <- socketConnection(host=host,port=port,blocking=TRUE,open="r+")  
  cat(cmd,file=con)
  response <- readLines(con)
  close(con)
  
  # trim header, split fields and convert results
  response <- response[2:length(response)]
  response <- laply(response,function(n) {
    sapply(strsplit(n,"|",fixed=TRUE),trim)
  })  
  response <- adply(response,c(1))
  response <- subset(response, select = -c(X1) )
  names(response) <- c("Peer.AS","IP","BGP.Prefix","CC",
                       "Registry","Allocated","Peer.AS.Name")
  return(response)
  
}
```



```{r}
# HELPER FUNCTION MENTIONED IN THE BOOK
# BUT NOT IN THE PRINTED LISTINGS

BulkOriginASN <- function(asn.list,host="v4.whois.cymru.com",port=43) {
  
  # setup query
  cmd <- "begin\nverbose\n" 
  ips <- paste(unlist(asn.list), collapse="\n")
  cmd <- sprintf("%s%s\nend\n",cmd,ips)
  
  # setup connection and post query
  con <- socketConnection(host=host,port=port,blocking=TRUE,open="r+")  
  cat(cmd,file=con)
  response <- readLines(con)
  close(con)
  
  # trim header, split fields and convert results
  
  response <- response[2:length(response)]
  response <- laply(response,.fun=function(n) {
    sapply(strsplit(n,"|",fixed=TRUE),trim)
  })
  
  response <- adply(response,c(1))
  response <- subset(response, select = -c(X1) )
  names(response) <- c("AS","CC","Registry","Allocated","AS.Name")
  
  return(response)
  
}
```





```{r}
# load the zeus botnet data used to perform the
# remainder of the analyses in the chapter
zeus <- read.table("data/zeus-book.csv", skip=5, header=FALSE, 
                   col.names=c("IP"))
ips <- as.character(zeus$IP) 
# get BGP origin data & peer data; 
origin <- BulkOrigin(ips)
g <- graph.empty() # start graphing
# Make IP vertices; IP endpoints are red
g <- g + vertices(ips, size=5, color=set2[4], group=1)
# Make BGP vertices
g <- g + vertices(origin$CC, size=5, color=set2[2], group=2)
# for each IP address, get the origin AS CC and return 
# them as a pair to create the IP->CC edge list
ip.cc.edges <- lapply(ips, function(x) {
  iCC <- origin[origin$IP==x, ]$CC
  lapply(iCC, function(y){
    c(x, y)
  })
})
```




```{r}
g <- g + edges(unlist(ip.cc.edges)) # build CC->IP edges
# simplify the graph by combining commmon edges
g <- simplify(g, edge.attr.comb=list(weight="sum"))
# delete any standalone vertices (lone wolf ASNs). In "graph" terms
# delete any vertex with a degree of 0
g <- delete.vertices(g, which(degree(g) < 1))
E(g)$arrow.size <- 0 # we hate arrows
# blank out all the IP addresses to focus on ASNs 
V(g)[grep("\\.", V(g)$name)]$name <- ""
```



```{r}
# Visualizing the ZeuS blocklist country cluster graph
# requires packages: igraph, plyr
# requires all objects from Listing 4-11
# this is a great layout for moderately sized networks. you can
# tweak the "n=10000" if this runs too slowly for you. The more
# iterations, the cleaner the graph will look
L <- layout.fruchterman.reingold(g, niter=10000, area=30*vcount(g)^2)
# plot the graph
par(bg = 'white', mfrow=c(1,1))
plot(g, margin=0, layout=L, vertex.label.dist=0.5, 
     vertex.label.cex=0.75, 
     vertex.label.color="black",
     vertex.label.family="sans", 
     vertex.label.font=2,
     main="ZeuS botnet nodes clustered by country")
```

Discuss in detail both the output and the syntax related to the task performed above.

```{r}
# Listing 4-14
# require package: igraph (4-11)
# requires object: V() (4-11), g (4-11)
# read in country code to name translation table
zeus.cc <- grep("[A-Z]", V(g)$name, value=TRUE)
zeus.cc <- zeus.cc[order(zeus.cc)]
# read in the country codes data frame
cc.df <- read.csv("data/countrycode_data.csv")
# display cc & name for just the ones from our data set
print(head(cc.df[cc.df$iso2c %in% zeus.cc, c(7,1)], n=10),
      row.names=FALSE)
```



```{r}
# requires objects: BulkOrigin() & BulkPeer() from book's web site &
#   set2 (4-4)
# create connected network of ZeuS IPs, ASNs, and ASN peers
# generates Figure 4-8
# require package: igraph, RColorBrewer
g <- graph.empty()
g <- g + vertices(ips, size=3, color=set2[4], group=1)
origin <- BulkOrigin(ips)
peers <- BulkPeer(ips)
# add ASN origin & peer vertices
g <- g + vertices(unique(c(peers$Peer.AS, origin$AS)),
                  size=3, color=set2[2], group=2)
# build IP->BGP edge list
ip.edges <- lapply(ips, function(x) {
  iAS <- origin[origin$IP==x, ]$AS
  lapply(iAS,function(y){
    c(x, y)
  })
})
```




```{r}
bgp.edges <- lapply(
  grep("NA",unique(origin$BGP.Prefix),value=TRUE,invert=TRUE),
  function(x) {
    startAS <- unique(origin[origin$BGP.Prefix==x,]$AS)
    lapply(startAS,function(z) {
      pAS <- peers[peers$BGP.Prefix==x,]$Peer.AS
      lapply(pAS,function(y) {
        c(z,y)
      })
    })
  })
```



```{r}
g <- g + edges(unlist(ip.edges))
g <- g + edges(unlist(bgp.edges))
g <- delete.vertices(g, which(degree(g) < 1))
g <- simplify(g, edge.attr.comb=list(weight="sum"))
E(g)$arrow.size <- 0
V(g)[grep("\\.", V(g)$name)]$name = ""
L <- layout.fruchterman.reingold(g, niter=10000, area=30*vcount(g)^2)
par(bg = 'white')
plot(g, margin=0, layout=L, vertex.label.dist=0.5, 
     vertex.label=NA,
     main="ZeuS botnet ASN+Peer Network")
```



Explain both the output and the syntax of the task completed above.




